---
title: 2023 Master's Thesis
author: Michael Corwin Getty

execute:
  echo: false
  
bibliography: references.bib
---

> **Abstract.** This is a suggested formatting for your abstract. This **quarto** document is a template for your report. It includes some basic ways of working with such documents. You are invited to learn more about developing online reports with this technology. 
>
> **Keywords.** bsvars, impulse responses, quarto, R, tax shocks

# Introduction

A chief concern in many parts of the world today is an ongoing housing crisis. House prices have been skyrocketing and home ownership decreasing. Recently in the United States there have been a few observable trends as a possible cause behind this, increasing use of property as an investment vehicle, depressed wages, and high disparities in intergenerational wealth.

The purpose of this paper is to analyze what portion of the change in homeownership and expenses are a result of these events, and which are more long term trends tied to predictable causes such as inflation and changing demographics. In pursuit of this I will utilize a Blinder-Oaxca decomposition to identify the source of the changes in housing expenditure. Additionally, I will plot the change in the share of renters vs owners in income demographics over time.

Due to data limitations, this study is carried out at the county level using all available data from the American Community Survey (ACS), utilizing census data.

The expected result is that there will be a gradual decrease in the rate of home-ownership over time, especially in the lowest income regions.


```{r}

#Load in packages to be used for future

library(fredr)
library(zoo)
library(dplyr)
library(coda)
library(HDInterval)
library(MASS)
library(distributionsrd)
library(arrayhelpers)
library(ggplot2)
library(qpcR)
library(DescTools)
library(tidyr)
library(tseries)
library(stargazer)
library(lmtest)
library(sandwich)
library(rcompanion)
library(car)
library(lmtest)
library(foreign)
library(systemfit)
library(xlsx)

#set key for accessing FRED data
fredr_set_key("f1d6de070cd07cd6028872a3bc573657")

#Load in some FRED data to start as a base analysis

HAI = fredr(series_id = "FIXHAI")
housepriceindex = fredr(series_id = "USSTHPI")

```

#REPEAT OF THE ITNRO

Recently the United States have been plagued by concerns of an ongoing housing crisis. There have been many proposed causes for it, housing being used as an investment vehicle in the wake of uncertainty stemming from the COVID-19 pandemic, depressed wages failing to keep up with rising expenses, intergenerational wealth gaps leading to disparity in opportunity. Each of these, perhaps, tells a part of the story.

What this paper seeks to analyze is how these recent trends have deviated from the trends of the past. Whether this is a new phenomenon or if it is in line with what would be expected were the economy not disrupted by emergency measures during the COVID pandemic. If there is a break in the trend, the question then becomes how has it broken?

In line with this objective, several facets of the housing market will be examined. As data is only publicly accessible at the county level, that is the level at which the analysis will be conducted. First and foremost, changes in both the stock of housing and how ownership rates have varied across counties will be of paramount importance. Of particular interest is whether certain counties have been impacted in an unusual manner, such as if rural areas are less impacted by any recent changes than metropolitan areas, or if any fluctuations are confined primarily to particular regions. 

The expected outcome is in line with what has been popularly observed in the media as of late, that home ownership rates are falling and that this is especially concentrated among lower income individuals. Due to data limitations, this paper is ill-suited to answering the impact on an individual level, but it can be indicative for if lower income regions are being more adversely affected. It is also capable of monitoring general trends in housing prices based on aggregate mortgage data. 


```{r}

#function initialization 


#data filtering
clean_data = function(datax, filterp = 1){
  
  
}

#Set up function for graph by quantile
set.seed(123456)
datax = matrix(rnorm(10000, 0, 1), nrow=10)



quants = 5
i=1
#tested, scales to any matrix data input, qunatile count, and auto resizes to fit

#data input, quantiles, name of graph
quant_graph = function(datax, quants, graph_name){
  n = nrow(datax)
  tiles = matrix(0,n,quants+1)
  for(i in 1:n){
  tiles[i,] = quantile(datax[i,], probs = seq(0,1, 1/quants))
  }
  m = ncol(tiles)-1
  #maybe can adjust color pallete to have variable input
  plotcol = colorRampPalette(c("red","blue"))(m+1)
  plot(tiles[,1], type = "b", ylim = c(min(datax),max(datax)), xlab = "Time", ylab = "Value", main = graph_name, col = plotcol[1])
  for (i in 1:m) {
    lines(tiles[,i+1], type = "b", col = plotcol[i+1])
  }
  #return quantiles incase it is needed elsewhere
  return(tiles)
}

#modified function for multiple trends over time, e.g. %rent vs homeowner
#filterp
# 1 = monthly
# 2 = quarterly
# 3 = yearly
# 4 = decade
trend_graph = function(datax, graph_name, filterp = 3, lower_bound = min(datax), upper_bound = max(datax)){
  case_when(
  filterp == 1 | datax$Date == as.yearmon(datax$Date),
  filterp == 2 | datax$Date == as.yearqtr(datax$Date),
  filterp == 3 | datax$Date == as.year(datax$Date)    
  
    )
  m = ncol(datax)-1
  #maybe can adjust color pallete to have variable input
  plotcol = colorRampPalette(c("red","blue"))(m+1)
  plot(tiles[,1], type = "b", ylim = c(lower_bound(datax),upper_bound), xlab = "Time", ylab = "Value", main = graph_name, col = plotcol[1])
  for (i in 1:m) {
    lines(tiles[,i+1], type = "b", col = plotcol[i+1])
  }
  
}


```

execute:
  echo: true
```
## References {.unnumbered}