---
title: 2023 Master's Thesis
author: Michael Corwin Getty

execute:
  echo: false
  
bibliography: references.bib
---

> **Abstract.** This is a suggested formatting for your abstract. This **quarto** document is a template for your report. It includes some basic ways of working with such documents. You are invited to learn more about developing online reports with this technology. 
>
> **Keywords.** bsvars, impulse responses, quarto, R, tax shocks

# Introduction

A chief concern in many parts of the world today is an ongoing housing crisis. House prices have been skyrocketing and home ownership decreasing. Recently in the United States there have been a few observable trends as a possible cause behind this, increasing use of property as an investment vehicle, depressed wages, and high disparities in intergenerational wealth.

The purpose of this paper is to analyze what portion of the change in homeownership and expenses are a result of these events, and which are more long term trends tied to predictable causes such as inflation and changing demographics. In pursuit of this I will utilize a Blinder-Oaxca decomposition to identify the source of the changes in housing expenditure. Additionally, I will plot the change in the share of renters vs owners in income demographics over time. 


```{r}

#Load in packages to be used for future

library(fredr)
library(zoo)
library(dplyr)
library(coda)
library(HDInterval)
library(MASS)
library(distributionsrd)
library(arrayhelpers)
#set key for accessing FRED data
fredr_set_key("f1d6de070cd07cd6028872a3bc573657")

#Load in some FRED data to start as a base analysis

HAI = fredr(series_id = "FIXHAI")
housepriceindex = fredr(series_id = "USSTHPI")

```

```{r}

#function initialization 


#data filtering
clean_data = function(datax, filterp = 1){
  
  
}

#Set up function for graph by quantile
set.seed(123456)
datax = matrix(rnorm(10000, 0, 1), nrow=10)



quants = 5
i=1
#tested, scales to any matrix data input, qunatile count, and auto resizes to fit

#data input, quantiles, name of graph
quant_graph = function(datax, quants, graph_name){
  n = nrow(datax)
  tiles = matrix(0,n,quants+1)
  for(i in 1:n){
  tiles[i,] = quantile(datax[i,], probs = seq(0,1, 1/quants))
  }
  m = ncol(tiles)-1
  #maybe can adjust color pallete to have variable input
  plotcol = colorRampPalette(c("red","blue"))(m+1)
  plot(tiles[,1], type = "b", ylim = c(min(datax),max(datax)), xlab = "Time", ylab = "Value", main = graph_name, col = plotcol[1])
  for (i in 1:m) {
    lines(tiles[,i+1], type = "b", col = plotcol[i+1])
  }
  #return quantiles incase it is needed elsewhere
  return(tiles)
}

#modified function for multiple trends over time, e.g. %rent vs homeowner
#filterp
# 1 = monthly
# 2 = quarterly
# 3 = yearly
# 4 = decade
trend_graph = function(datax, graph_name, filterp = 3, lower_bound = min(datax), upper_bound = max(datax)){
  case_when(
  filterp == 1 | datax$Date == as.yearmon(datax$Date),
  filterp == 2 | datax$Date == as.yearqtr(datax$Date),
  filterp == 3 | datax$Date == as.year(datax$Date)    
  
    )
  m = ncol(datax)-1
  #maybe can adjust color pallete to have variable input
  plotcol = colorRampPalette(c("red","blue"))(m+1)
  plot(tiles[,1], type = "b", ylim = c(lower_bound(datax),upper_bound), xlab = "Time", ylab = "Value", main = graph_name, col = plotcol[1])
  for (i in 1:m) {
    lines(tiles[,i+1], type = "b", col = plotcol[i+1])
  }
  
}


```

execute:
  echo: true
```
## References {.unnumbered}